---
title: "Class 11: Structural Bioinformatics"
author: "Jaya George"
date: "5/7/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB Database

The [PDB](http://www.rcsb.org/) is the main repository for biomolecular structure data. 

Here we examine the contents of the PDB:
```{r}
db <- read.csv("Data Export Summary.csv", row.names = 1)
head(db)
```

How many are X-ray, etc.
```{r}
(db$Total/ sum(db$Total)) * 100 
```

How many are Protein
```{r}
(sum(db$Proteins)/sum(db$Total))*100
```

We see that it is predominantly X-ray crystallography and protein. 

Download through datapast (add-in past as data.frame)
```{r}
library(datapasta)

tmp <- data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "NMR", "Electron Microscopy", "Other",
                           "Multi Method", "Total"),
              Proteins = c(126880, 11062, 2277, 256, 129, 140604),
         Nucleic.Acids = c(2012, 1279, 31, 4, 5, 3331),
    ProteinComplex = c(6547, 259, 800, 6, 2, 7614),
                 Other = c(8, 8, 0, 13, 1, 30),
                 Total = c(135447, 12608, 3108, 279, 137, 151579)
)
```

Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?
There are 1157 as of 2019-05-07. See: http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=F0CA5320

##Section 3 Using bio3d

```{r}
library(bio3d)

pdb <- read.pdb("1hsg.pdb")
pdb
```

```{r}
attributes(pdb)
```

```{r}
head(pdb$atom)
```

Print a subset of $atom data for the first two atoms
```{r}
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
```

Note that individual $atom records can also be accessed like this
```{r}
pdb$atom$elety[1:2]
```

Which allows us to do the following
```{r}
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab= "B-factor")
```

Print a summary of the coordinate data in $xyz
```{r}
pdb$xyz
```

Examine the row and column dimensions
```{r}
dim(pdb$xyz)
```

Print coordinates for the first two atom
```{r}
pdb$xyz[ 1, atom2xyz(1:2) ]
```

Atom selection is done via the function **atom.select()**


Select all C-alpha atoms (return their indices)
```{r}
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```

Print details of the first few selected atoms
```{r}
head( pdb$atom[ca.inds$atom, ] )
```

And selected xyz coordinates
```{r}
head( pdb$xyz[, ca.inds$xyz] )
```

Select chain A
```{r}
a.inds <- atom.select(pdb, chain="A")
```

Select C-alphas of chain A
```{r}
ca.inds <- atom.select(pdb, "calpha", chain="A")
```

We can combine multiple selection criteria to return their
intersection
```{r}
cab.inds <- atom.select(pdb, elety=c("CA","CB"), chain="A",
resno=10:20)
```

Write.pdb function
```{r}
prot.pdb <- atom.select(pdb, "protein", value= TRUE)
write.pdb(prot.pdb, file="1hsg_protein.pdb")
```

```{r}
ligand.pdb <- atom.select(pdb, "ligand", value= TRUE)
write.pdb(ligand.pdb, file="1hsg_ligand.pdb")
```

##Section 5

```{r}
aa <- get.seq("1ake_A")
```

Blast or hmmer search 
```{r}
b <- blast.pdb(aa)
```

Plot a summary of search results
```{r}
hits <- plot(b)
```
Blast results. Visualize and filter blast results through function plot.blast(). Here we proceed with only the top scoring hits (black).

```{r}
head(hits$pdb.id)
```

Fetch PDBs
```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

Align structures
```{r}
pdbs <- pdbaln(files)
```

Vector containing PDB codes
```{r}
ids <- basename.pdb(pdbs$id)
```

Draw schematic alignment
```{r}
plot(pdbs, labels=ids)
```
Plot is the schematic representation of alignment. Grey regions depict aligned residues, while white depict gap regions. The red bar at the top depict sequence conservation.

##Sequence Conservation Analysis

Calculate sequence conservation
```{r}
cons <- conserv(pdbs, method="entropy22")
```

SSE annotations
```{r}
sse <- pdbs2sse(pdbs, ind=1, rm.gaps=FALSE)
```

Plot conservation per residue
```{r}
plotb3(cons, sse=sse, ylab="Sequence entropy")
```
Sequence conservation per residue. Here, Shannon's information entropy is used tomeasure the diversity per alignment column. SSEs are depicted with dark (helices) and light (sheets) grey boxes in marginal regions.

##Annotate collected PDB structures

Function pdb.annotate() provides a convenient way of annotating the PDB files we have collected. Below we use the function to annotate each structure to its source species. 
```{r}
anno <- pdb.annotate(ids)
print(unique(anno$source))
```






























