---
title: "Class 8: Machine Learning "
author: "Jaya George"
date: "4/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#K-means clustering
Let's start with an example of running the **kmeans()** function

```{r}
#Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

Use the kmeans() function setting k to 2 and nstart=20

Inspect/print the results

>Q. How many points are in each cluster? 
    30

>Q. What ‘component’ of your result object details (find answers in value section of ?kmeans)
 - cluster size?
 - cluster assignment/membership?
 - cluster center?

```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```

```{r}
km$cluster
```

```{r}
km$centers
```

>Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
```{r}
plot(x, col=km$cluster)
points(km$centers, pch=18, col="blue", cex=3)
```
 

## Hierarchical Clustering example

We must give the hclust() fucntion a distance matrix not the raw data as an input

```{r}
#Distance matrix
d <- dist(x)

#Clustering
hc <- hclust(d)
plot(hc)
```

```{r}
plot(hc)
abline(h=6, col="red")
```

```{r}
cutree(hc, h=6)
```

```{r}
cutree(hc, k=2)
```

```{r}
cutree(hc, k=3)
```


Another example but with more real life clustering

# Step 1. Generate some example data for clustering
```{r}
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```

# Step 2. Plot the data without clustering

```{r}
plot(x)
```

# Step 3. Generate colors for known clusters (just so we can compare to hclust results)

```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

```{r}
table(col)
```

```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
abline(h=2.5, col="red")
abline(h=2.2, col="blue")
```

```{r}
gp2 <- cutree(hc, k=2)
gp2
```

```{r}
gp3 <- cutree(hc, k=3)
gp3
```

```{r}
plot(x, col= gp2)
```

```{r}
plot(x, col= gp3)
```


```{r}
table(gp2)
```

```{r}
table(gp3)
```

```{r}
table(gp2, gp3)
```



##PCA: Pricipal Component Analysis

We will use the **prcomp()** function for PCA

```{r}
#Example Data
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1) 
head(mydata) #head by default just shows the top 6 rows
```

```{r}
nrow(mydata)
#100 genes
```

```{r}
ncol(mydata)
#10 variations per gene
```

```{r}
colnames(mydata)
```


#Running our PCA on the transpose of our data
```{r}
pca <- prcomp(t(mydata), scale=TRUE)
#t switches row and column
pca
```

PCA plot of PC1 vs. PC2 2-D plot
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2") 
```

Calculate the percent variance captures in each PC
```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
head(pca.var.per)
```

```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```

# A vector of colors for wt and ko samples
```{r}
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
```


#PCA of UK food data

```{r}
x <- read.csv("class08.csv", row.names=1)
head(x)
```

```{r}
pca <- prcomp( t(x))
summary(pca)
```

# Plot PC1 vs PC2
```{r}
mycols <- c("yellow", "red", "blue", "darkgreen")
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col=mycols)
```

How much variation in the original data each PC accounts for:
```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

or second row here
```{r}
z <- summary(pca)
z$importance
```

Plot of the variances (eigenvalues) with respect to the principal component number (eigenvector number)
```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

##Variable loadings: Lets focus on PC1 as it accounts for > 90% of variance 
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```


