---
title: "Class 12: Structural Bioinformatics (Pt 2)"
author: "Jaya George"
date: "5/9/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Setup for Docking
We will first prepare our HIV-Pr system for drug docking by making a protein only PDB format file (ie. )

 Obtaining and inspecting our input structure.
```{r}
library(bio3d)
file <- get.pdb("1hsg")
```

```{r}
pdb <- read.pdb(file)
pdb
```

Prepare initial protein and ligand input files
```{r}
prot <- atom.select(pdb, "protein", value = TRUE)
write.pdb(prot, file= "1hsg_protein.pdb")
prot
```

```{r}
lig <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(lig, file= "1hsg_ligand.pdb")
lig
```

We will load these into ADT to add hydrogens and make PDBQT files for each

No we run autodoc vina with the cmd

Inspecting your docking results 
```{r}
library(bio3d)
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

To assess the results quantitatively we will calculate the RMSD (root mean square distance) between each of the docking results and the known crystal structure using the bio3d package.
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```


##Exploring the conformational dynamics of proteins

 Normal Mode Analysis (NMA) for flexibility prediction
 
 Example structure of hen egg white lysozyme (PDB id 1hel)
```{r}
library(bio3d)
pdb <- read.pdb("1HEL")
modes <- nma(pdb)
plot(modes, sse=pdb)
```
 
Visualize NMA results
```{r}
m7 <- mktrj(modes, mode=7, file="nma_7.pdb")
m7
```













