---
title: "HW06"
author: "Jaya George"
date: "4/18/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Original Code
```{r}
library(bio3d)

s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

##Cluster Dendrogram
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc, xlab= "B factor Trends")
```

##Function

```{r}
pdb.analysis <- function(pdb.name) {
  pdb_file <- read.pdb(pdb.name)
  chainA <- trim.pdb(pdb_file, chain="A", elety= "CA" )
  bfactor <- chainA$atom$b
  plotb3(bfactor, sse=chainA, typ="l", ylab="Bfactor")
}
```

```{r}
pdb.analysis("4AKE")
```

```{r}
pdb.analysis("1AKE")
```

```{r}
pdb.analysis("1E4Y")
```

####When inputting a pdb.name into pdb.analysis function it must be the given 4 character code for a specific protein in quotes.

####The function allows for one to analyze a selected protein. Specifically the protein's coordinate file is read on the Protein Data Bank through read.pdb. Next a smaller PDB object is produced through trim.pdb, in this case it is Chain A. Next it is specified further that atom B is to be analyzed. 

####The output of this function is a standard scatter plot with the secondary structure of the proteins in the marginal regions. The plots compare the Bfactor of each protein to its residue. 
