---
title: "Class 7: R functions and packages"
author: "Jaya George"
date: "4/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Functions revisited

We will source a file with our functions from last day

```{r}
source("http://tinyurl.com/rescale-R")
```

Try out the last day's rescale() function

```{r}
rescale(1:10)
```

Try rescale2() function that catches string inputs

```{r}
#rescale2(c(1:10, "string"))
```

#Finding missing NA values in vectors
Start with a simple example of a larger problem I am trying to solve

```{r}
x <- c(1,2,NA,3,NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
```

```{r}
is.na(y)
```

Try putting these together with an AND (TRUE= position with both NA)

```{r}
is.na(x) & is.na(y)
```

Take sum() to find out how many TRUE values we have and thus how many NAs we have in both x and y (because TRUE=1 and FALSE=0)

```{r}
sum(is.na(x) & is.na(y))
```

Now I can make this into our first function...

```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x, c(NA, 3, NA, 2, NA))
```

Test, test, test (When the length of the vectors aren't the same the first value of the shorter vector is added to the end -> recylcing)
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

```

```{r}
both_na(x, y2)
```

```{r}
y3 <- c( 1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```

```{r}
3==3
```

```{r}
3 != 2
```

```{r}
length(x) != length(y2)
```

Now lets try both_na2() function on our different length input vectors

```{r}
both_na2 <- function(x, y) {
 if(length(x) != length(y)) {
 stop("Input x and y should be the same length")
 }
 sum( is.na(x) & is.na(y) )
}
```

```{r}
#both_na2 (x, y2)
```

```{r}
which( c(F, F, T, F, T))
```

```{r}
which(is.na ( c(1,2,NA,4) ))
```

```{r}
both_na3 <- function(x, y) {
 if(length(x) != length(y)) {
 stop("Input x and y should be vectors of the same length")
 }

 na.in.both <- ( is.na(x) & is.na(y) )
 na.number <- sum(na.in.both)
 na.which <- which(na.in.both)
 message("Found ", na.number, " NA's at position(s):",
 paste(na.which, collapse=", ") )

 return( list(number=na.number, which=na.which) )
}
```

```{r}
x <- c(1,2,NA,3,NA)
y <- c(NA, 3, NA, 3, 4)
both_na3(x,y)
```

Simplify into single vectors

```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```

```{r}
intersect(x,y)
```

```{r}
x %in% y
```

```{r}
which( x%in%y )
```

```{r}
x[ x%in%y ]
```

```{r}
y[ y %in% x]
```

A working snippet!

```{r}
cbind( x[ x %in% y ], 
       y[ y %in% x ] )
```

Use R studio shortcut 'Code-> Extract function' to turn our snippet into a working function
```{r}
gene_intersect <- function(x, y) {
  cbind( x[ x %in% y ], 
         y[ y %in% x ] )
}
```

```{r}
gene_intersect(df1$IDs, df2$IDs)
```

Our input $IDs column name may change:add flexibility by allowing the user to specify the gene containing column name

```{r}
gene_intersect2 <- function(df1, df2) {
 cbind( df1[ df1$IDs %in% df2$IDs, ],
 df2[ df2$IDs %in% df1$IDs, "exp"] )
}
```

```{r}
gene_intersect2(df1, df2)
```

Change column names to something nicer

```{r}
gene_intersect3 <- function(df1, df2, gene.colname="IDs") {

 cbind( df1[ df1[,gene.colname] %in%
 df2[,gene.colname], ],
 exp2=df2[ df2[,gene.colname] %in%
 df1[,gene.colname], "exp"] )

}
```

```{r}
gene_intersect3(df1, df2)
```

```{r}
merge(df1, df2, by="IDs")
```

Grade function -> calculate mean score and drop the lowest single score

```{r}
x <- c( 100, 90, 90, 90, 90, 90, 97, 80)
```

```{r}
grade <- function(x) {
  (sum(x)- min(x))/ (length(x)-1)
}
```

```{r}
grade(x)
```




